---
title: "UMAP and Cluster marker identification for `r params$dataset_name`"
author:
  - "trnguyen@ukaachen.de"
date: "Last update on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 3
    number_sections: false
    theme: lumen
params:
  path.to.input.sobj: NA
  path.to.wd: NA
  dataset_name: NA
  path.to.output : NA
---

```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```


<style type="text/css">
    div.datatables { height: auto !important;}
</style>

# General information

- Path to the original input Seurat R.object before removing clusters `r clusters.to.be.removed`: `r file.path(path.to.input, name.of.s.obj)`

- Path to the Seurat R.object after removing clusters `r clusters.to.be.removed`: `r file.path(path.to.output, sprintf("%s.%s", status, save.obj.name))`

- Input data R.object: `r basename(params$path.to.input.sobj)`

- Removed clusters: `r clusters.to.be.removed`

- Re-integrating data from multiple samples after removing clusters: `r with.re.integration`




```{r echo=FALSE, warning=FALSE, results='hide', message=FALSE}
#####----------------------------------------------------------------------#####
#
# 02: VDJ data analysis on the 2nd dataset, focus on YFP cells and non YFP cells
#
# trnguyen@ukaachen.de
#####----------------------------------------------------------------------#####

path.to.wd <- params$path.to.wd
path.to.input.sobj <- params$path.to.input.sobj

source(file.path(path.to.wd, "00_import_libraries.R"))
source(file.path(path.to.wd, "00_helper_functions.R"))

s.obj <- readRDS(path.to.input.sobj)
#####----------------------------------------------------------------------#####
# CONFIGURATIONS AND PREPARATION
#####----------------------------------------------------------------------#####
```


```{r echo=FALSE, warning=FALSE, results='asis', message=FALSE, fig.width=14, fig.height=10, dev='svg'}
if ("INTE_UMAP" %in% names(s.obj@reductions)){
  p <- DimPlot(object = s.obj, reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, pt.size = 1)  
} else {
  p <- DimPlot(object = s.obj, reduction = "RNA_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, pt.size = 1)
}
p
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
cluster.markers <- FindAllMarkers(object = s.obj, assay = "RNA", test.use = "wilcox")
cluster.markers.raw <- cluster.markers

cluster.markers <- subset(cluster.markers, cluster.markers$p_val_adj <= 0.05 & cluster.markers$avg_log2FC >= 0)
```

## Tables of all cluster marker genes {.tabset}
```{r echo=FALSE, include=FALSE}
cluster.markers %>% create_dt()
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}

for (cluster.id in sort(unique(cluster.markers$cluster))){
  tmp.cluster.markers <- subset(cluster.markers, cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  dir.create(file.path(path.to.output, "raw"), showWarnings = FALSE, recursive = TRUE)
  dir.create(file.path(path.to.output, "up_DEGs"), showWarnings = FALSE, recursive = TRUE)
  
  writexl::write_xlsx(x = tmp.cluster.markers, 
                    file.path(path.to.output, "up_DEGs",
                              sprintf("%s.%s", status, str_replace(save.obj.name, ".rds",
                                                                   sprintf("%s.xlsx", cluster.id)))))
  
  writexl::write_xlsx(x = subset(cluster.markers.raw, cluster.markers.raw$cluster == cluster.id) %>% arrange(desc(avg_log2FC)), 
                    file.path(path.to.output, "raw",
                              sprintf("%s.%s", status, str_replace(save.obj.name, ".rds",
                                                                   sprintf("%s.raw.xlsx", cluster.id)))))

  tmp.cluster.markers <- tmp.cluster.markers %>% mutate_if(is.numeric, round, 6) 
  
  cat(paste("\n\n### Cluster: ", cluster.id, "##\n"))
  print( htmltools::tagList(datatable(tmp.cluster.markers, extensions = 'Buttons',
                filter = "top",
                options = list(dom = 'Blfrtip',
                               buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                               lengthMenu = list(c(10,25,50,-1),
                                                 c(10,25,50,"All")),
                               columnDefs = list(list(
                                 targets = "_all",
                                 render = JS(
                                   "function(data, type, row, meta) {",
                                   "return type === 'display' && data != null && data.length > 100 ?",
                                   "'<span title=\"' + data + '\">' + data.substr(0, 100) + '...</span>' : data;",
                                   "}")
                               ))
                ))))
  cat("\n \n")
}

```


## Top-12 cluster marker genes `FeaturePlot` {.tabset}
Here we just show the top-12 genes (ranked by `avg_log2FC`) for each cluster. 

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
for (cluster.id in sort(unique(s.obj@meta.data$seurat_clusters))){
  DefaultAssay(s.obj) <- "RNA"
  cat(sprintf("### Cluster %s \n", cluster.id))
  tmp.cluster.markers <- subset(cluster.markers, cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  p <- FeaturePlot(object = s.obj, reduction = "INTE_UMAP", features = head(tmp.cluster.markers, 12)$gene, ncol = 3, label = TRUE, pt.size = 0.5, label.size = 5, label.color = "red")  
  print(p)
  cat("\n \n")
}
```

## Top-12 cluster marker genes `Violin Plot` {.tabset}
Here we just show the top-12 genes (ranked by `avg_log2FC`) for each cluster. 

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=20, fig.height=12}
for (cluster.id in sort(unique(s.obj@meta.data$seurat_clusters))){
  DefaultAssay(s.obj) <- "RNA"
  cat(sprintf("### Cluster %s \n", cluster.id))
  tmp.cluster.markers <- subset(cluster.markers, cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  p <- VlnPlot(object = s.obj, features = head(tmp.cluster.markers, 12)$gene)  
  print(p)
  cat("\n \n")
}
```
